Supplement to ``{\mytitle}''

### 2022-10-17 ms.Rnw
Updated how Model 1 is replicated.

# Setting algorithmic hyper-parameters based on RUN_LEVEL. 
# 
#                    RUN_LEVEL =  1,   2,    3
N_PARAMS   <- switch(RUN_LEVEL,  10,  20,  300)  # Number of parameters to sample
N_LEE_SIMS <- switch(RUN_LEVEL,  10,  20,   20)  # Number of simulations per parameter
NMIF       <- switch(RUN_LEVEL,   5,  20,   50)  # Number of MIF iterations
NP_MIF     <- switch(RUN_LEVEL,  20,  50,  100)  # Number of particles for IF2. This is too few for proper optimization, but matches Lee et al
NP_EVAL    <- switch(RUN_LEVEL,  50, 200, 2000)  # Number of particles used to evaluate the likelihood 
NREPS_EVAL <- switch(RUN_LEVEL,   3,   5,   36)  # Number of replicates of likelihood evaluation.

@@

# Get starting parameters close to those that are provided
# in Lee et al (2020)
starting_params <- get_h1_startVals(N_PARAMS)

# Create epi model for haiti1
h1_epi <- haiti1()

# Do a small global search for starting parameters for beta2-beta6
bounds <- tribble(
  ~param, ~lower, ~upper,
  "beta2", 1e-08, 10,
  "beta3", 1e-08, 10,
  "beta4", 1e-08, 10,
  "beta5", 1e-08, 10,
  "beta6", 1e-08, 10
)

lower <- bounds$lower
names(lower) <- bounds$param

upper <- bounds$upper
names(upper) <- bounds$param

# Get a sequence of starting values for the betas
betas <- sobol_design(
  lower = lower,
  upper = upper,
  nseq  = N_PARAMS
)

# Combine betas with the starting_params data.frame object.
epi_params <- cbind(starting_params, betas)

# Get the names of all of the remaining coefficients (fixed parameters).
not_in_start_params <- names(coef(h1_epi))[!names(coef(h1_epi)) %in% colnames(epi_params)]

# Add fixed parameters to the matrix of starting parameters
for (i in 1:length(not_in_start_params)) {
  epi_params[, not_in_start_params[i]] <- coef(h1_epi)[not_in_start_params[i]]
}

# The process model of Lee et al (2020) didn't include over-dispersion
epi_params$sig_sq <- 0

# re-ordering starting parameters to match model order
epi_params <- epi_params[, names(coef(h1_epi))]

# Setting the random walk standard deviation. 
epi_rw <- rw.sd(
  # Parameters close to desired values
  beta1 = 0.001,
  tau   = 0.001,
  nu    = 0.001,
  rho   = 0.001,
  # Parameters not close to desired values
  beta2 = 0.02,
  beta3 = 0.02,
  beta4 = 0.02,
  beta5 = 0.02,
  beta6 = 0.02,
  E_0   = ivp(0.1),
  I_0   = ivp(0.1)
)

registerDoRNG(48963587)

# Check if RUN_LEVEL == 3 results exists, if so, use them.
# Include a flag to check if RUN_LEVEL == 3. If so, we might want to recompute.
if (RUN_LEVEL != 3 & file.exists("model1/run_level_3/lee1_epi_fit.rds")) {
  lee1_epi_mif <- readRDS("model1/run_level_3/lee1_epi_fit.rds")
} else {
  lee1_epi_mif <- bake(
    file = paste0("model1/", rl_dir, "lee1_epi_fit.rds"),
    {
      # Create a list to save all of the results.
      lee1_epi_mif_results <- list()

      # Loop through each of the starting parameters
      foreach(
        i = 1:nrow(epi_params),
        .packages = c('pomp'),
        .combine = c
      ) %dopar% {

        # Get the ith set of starting parameters
        r_params <- unlist(epi_params[i, ])

        # Set the coefficients of model 1 to the ith starting values
        coef(h1_epi) <- r_params

        # Perform short IF2 from the starting value
        mif2(
          h1_epi,
          Np = NP_MIF,
          Nmif = NMIF,
          cooling.fraction.50 = 0.5,
          rw.sd = epi_rw
        )
      } -> lee1_epi_mif_out  # Contains nrow(epi_params) mif2d objects

      # For each output of the MIF2 objects, simulate Model 1 with the IF2
      # parameters and save the states from the final week of the simulation
      foreach(i = 1:length(lee1_epi_mif_out), .combine = rbind) %do% {

        # Get IF2 parameter set
        guess <- coef(lee1_epi_mif_out[[i]])

        # Simulate with the parameter set
        sims <- simulate(
          h1_epi,
          n = N_LEE_SIMS,
          format = 'data.frame',
          params = guess
        )  %>%
          dplyr::filter(week == max(week)) %>%  # Just keep final states, to use later.
          dplyr::select(-.id, -week)

        sims$param_set <- i
        sims
      } -> start_states

      lee1_epi_mif_results$lee1_epi_mif_params <- t(coef(lee1_epi_mif_out))
      lee1_epi_mif_results$start_states <- start_states
      lee1_epi_mif_results
    }
  )
}

# Some clean-up of unused variables, to avoid "Environment clutter".
rm(epi_rw, epi_params, not_in_start_params,
   lower, upper, betas, bounds, starting_params, i)
gc()

@@

      # Create data.frame to save results
      lee_epi_ll <- data.frame(
        "pfLL" = rep(0, length(lee1_epi_mif)),
        "pfse" = rep(0, length(lee1_epi_mif))
      )

      for (j in 1:nrow(lee1_epi_mif$lee1_epi_mif_params)) {
        # Get parameters of interest
        pf_params <- lee1_epi_mif$lee1_epi_mif_params[j, ]

@@

# Our parameter values are now very similar to those used by Lee et al (2020). 
# We therefore use the same rw.sd that was used in their source code. 
end_rw <- rw.sd(
  beta1 = 0.05,
  tau   = 0.005,
  nu    = 0.04,
  rho   = 0.01,
  beta2 = 0.05,
  beta3 = 0.05,
  beta4 = 0.05,
  beta5 = 0.05,
  beta6 = 0.05
)

registerDoRNG(21489587)  # For reproducible computing

# Store the parameter values from the end of the IF2 search for the epidemic 
# model as the start values for the IF2 search for the endemic model, as was 
# described by Lee et al (2020)
end_start <- lee1_epi_mif$lee1_epi_mif_params

# Check if the run_level_3 results exist. If they do, load these results. 
# If not, compute the run_level_<RUN_LEVEL> results and save them. 
if (RUN_LEVEL != 3 & file.exists("model1/run_level_3/lee1_end_fit.rds")) {
  lee1_end_mif <- readRDS("model1/run_level_3/lee1_end_fit.rds")
} else {
  lee1_end_mif <- bake(
    file = paste0("model1/", rl_dir, "lee1_end_fit.rds"),
    {
      foreach(  # Parallel execution of the IF2 algorith. 
        i = 1:nrow(end_start),
        .packages = c('pomp'),
        .combine = rbind
      ) %dopar% {

        # Get the ith set of starting parameters for the endemic model. These 
        # correspond to the parameters resulting from the ith execution of the 
        # IF2 algorithm on the epidemic model
        r_params <- end_start[i, ]

        # Calculate the starting states for each set of IF2 parameters. Because 
        # there are <N_LEE_SIMS> simulations for each set of parameters, we 
        # take the median value for each set of parameters as the starting 
        # value for the endemic model. 
        lee1_epi_mif$start_states %>%
          filter(param_set == i) %>%
          select(-param_set) %>%
          sapply(., median, na.rm = TRUE) -> rinit_parms

        # Set the initial state by creating a new rinit function for endemic 
        # model. 
        new_rinit <- Csnippet(
          sprintf(
            "
          S = nearbyint(%f);
          E = nearbyint(%f);
          I = nearbyint(%f);
          A = nearbyint(%f);
          R = nearbyint(%f);
          incid = nearbyint(%f);
          foival = %f;
          Str0 = nearbyint(%f);
          Sout = nearbyint(%f);
          Sin = nearbyint(%f);
          ",
          rinit_parms['S'],
          rinit_parms['E'],
          rinit_parms['I'],
          rinit_parms['A'],
          rinit_parms['R'],
          rinit_parms['incid'],
          rinit_parms['foival'],
          rinit_parms['Str0'],
          rinit_parms['Sout'],
          rinit_parms['Sin']
          )
        )

        # Create a new pomp object representing the endemic model, where the 
        # starting state is the median of the latent states from the epidemic 
        # model for the parameter set i. 
        h1_end <- h1_end %>% pomp(
          rinit = new_rinit,
          statenames = c(
            "S", "E", "I", "A", "R", "incid",
            "foival", "Str0", "Sout", "Sin"
          )
        )

        # Set model coefficients to be the parameter set i.
        coef(h1_end) <- r_params
        mf_end <- mif2(
          h1_end,
          Np = NP_MIF,
          Nmif = NMIF,
          cooling.type = "hyperbolic",
          cooling.fraction.50 = 0.05,  # Matches Lee et al (2020)
          rw.sd = end_rw
        )

        coef(mf_end)
      }
    }
  )
}

rm(end_rw)
gc()

@@

      # Create data.frame to save results
      lee_end_ll <- data.frame(
        "pfLL" = rep(0, nrow(lee1_end_mif)),
        "pfse" = rep(0, nrow(lee1_end_mif))
      )

      # Perform particle filter to get likelihood estimates
      for (j in 1:nrow(lee_end_ll)) {
        # Get parameters of interest
        pf_params <- lee1_end_mif[j, ]

@@

rm(lee1_end_mif, lee1_epi_mif)
gc()

# MIF_params <- matrix(nrow = length(lee1_end_mif$m2) / 2, ncol = length(coef(lee1_end_mif[[1]])))
# all_sims_end <- data.frame()
# liks <- rep(0, nrow(end_start))
# for (i in 1:nrow(end_start)) {
#   MIF_params[i, ] <- coef(lee1_end_mif[[(2 * i) - 1]])
#   all_sims_end <- rbind(all_sims_end, lee1_end_mif[[(2 * i)]])
#   liks[i] <- logLik(lee1_end_mif[[(2 * i) - 1]])
# }
#
# colnames(MIF_params) <- names(coef(lee1_end_mif[[1]]))
# end_results <- as.data.frame(MIF_params)
# end_results$logLik <- liks

# GGally::ggpairs(
#   filter(end_results, logLik > -2200),
#   columns = c("rho", "tau", "beta1", "nu", "logLik")
# )

#-------------------------- Step 6 ----------------------

mod1_lee_ll <- max(lee_end_ll$pfLL, na.rm = TRUE) + max(lee_epi_ll$pfLL, na.rm = TRUE)

rm(
  lee_end_ll, lee_epi_ll, h1_end, h1_epi, end_start,
  N_LEE_SIMS, N_PARAMS, NMIF, NP_EVAL, NP_MIF, NREPS_EVAL
)
gc()
@

@@

\begin{tabular}{|c|c|c|c|}
\hline
 & \thead{Model~1} & \thead{Model~2} & \thead{Model~3}
\\
\hline
\hline
\multirow{2}{*}{Log-likelihood} &
  $\Sexpr{myround(mod1_ll, 1)}$ &
  $\Sexpr{myround(mod2_ll, 1)}$ &
  $\Sexpr{myround(mod3_spatPomp_res$ll, 1)}$ \\
    & ($\Sexpr{myround(mod1_lee_ll, 1)}$)\footnotemark[1] &
  ($\Sexpr{myround(h2_fit$leeFit_ll, 1)}$) &
  ($\Sexpr{myround(max(lee3_results[["all_ll_evals"]]), 1)}$)\footnotemark[2]
\\
\hline
Number of &
  $\Sexpr{as.character(mod1_n_params)}$ &
  $\Sexpr{as.character(mod2_n_params)}$ &
  $\Sexpr{mod3_spatPomp_n_params}$ \\
 Fit Parameters &
 (\Sexpr{as.character(mod1_lee_n_params)}) &
 ($\Sexpr{as.character(h2_fit$leeFit_n_params)}$) &
 ($\Sexpr{as.character(lee3_results$n_params, 1)}$)
\\
\hline
\multirow{2}{*}{AIC} &
  $\Sexpr{myround(mod1_aic, 1)}$ &
  $\Sexpr{myround(mod2_aic, 1)}$ &
  $\Sexpr{myround(mod3_spatPomp_aic, 1)}$ \\
  & ($\Sexpr{myround(mod1_lee_aic, 1)}$)\footnotemark[1] &
  ($\Sexpr{myround(mod2_lee_aic, 1)}$) &
  ($\Sexpr{myround(mod3_lee_coupled_aic, 1)}$)\footnotemark[2]
\\
\hline
\thead{Log-ARMA(2,1) \\ Log-likelihood} &
  $\Sexpr{myround(ARMA_benchmarks$m1, 1)}$ &
  $\Sexpr{myround(ARMA_benchmarks$m2, 1)}$ &
  $\Sexpr{myround(ARMA_benchmarks$m3, 1)}$
\\
\hline
\end{tabular}
\caption{\label{tab:likes}Log-likelihood values for each models compared to their ARMA benchmarks.
Values in parenthesis are corresponding values using \citet{lee20} parameter estimates.  $\textsuperscript{1}$The reported likelihood is an upper bound of the likelihood of the \citet{lee20} model as it is the largest likelihood obtained using their parameter calibration regime. $\textsuperscript{2}$\cite{lee20} fit Model~3 to a subset of the data (March 2014 onward, excluding data from Ouest in 2015-2016).
On this subset, their model has a likelihood of $\Sexpr{myround(max(lee3_results$sub_ll_evals), 1)}$.
On this same subset, our model has a likelihood of $\Sexpr{myround(mod3_spatPomp_res$subset_ll, 1)}$.
}
\end{table}

>>>>

# Setting algorithmic hyper-parameters based on RUN_LEVEL.
#
#                    RUN_LEVEL =  1,   2,    3
N_PARAMS   <- switch(RUN_LEVEL,  10,  20,  nrow(h1LeeStartsEpi))  # Number of parameters to sample
N_LEE_SIMS <- switch(RUN_LEVEL,  10,  20,   20)  # Number of simulations per parameter
NP_EVAL    <- switch(RUN_LEVEL,  50, 200, 2000)  # Number of particles used to evaluate the likelihood
NREPS_EVAL <- switch(RUN_LEVEL,   3,   5,   36)  # Number of replicates of likelihood evaluation.

h1_epi <- haiti1()

@@

@@

      which_params <- sample(
        1:nrow(h1LeeStartsEpi),
        size = N_PARAMS,
        replace = FALSE
      )
      epi_params <- h1LeeStartsEpi[which_params, ]

      # Create data.frame to save results
      lee_epi_ll <- data.frame(
        "pfLL" = NA_real_,
        "pfse" = NA_real_,
        "parid" = epi_params$parid
      )

      for (j in 1:nrow(lee_epi_ll)) {
        # Get parameters of interest
        pf_params <- epi_params[j, -1][names(coef(h1_epi))]

@@

@@

      end_params <- h1LeeStartsEnd[h1LeeStartsEnd$parid %in% lee_epi_ll$parid, ]

      # Create data.frame to save results
      lee_end_ll <- data.frame(
        "pfLL" = NA_real_,
        "pfse" = NA_real_,
        'parid' = end_params$parid
      )

      # Perform particle filter to get likelihood estimates
      for (j in 1:nrow(lee_end_ll)) {
        # Get parameters of interest
        pf_params <- end_params[j, ][names(coef(h1_end))]

@@

# Combine epidemic and endemic likelihoods by parid to get the likelihood of
# the complete model.
mod1_all_likes <- dplyr::inner_join(
  x = lee_epi_ll,
  y = lee_end_ll,
  by = "parid"
  ) %>%
  dplyr::mutate(
    joint_ll = pfLL.x + pfLL.y
  )

# Approximate the likelihood of the Lee et al (2020) version of Model 1 by
# using the parameter sets that resulted in the maximum likelihood.
mod1_lee_ll <- max(mod1_all_likes$joint_ll, na.rm = TRUE)

rm(
  lee_end_ll, lee_epi_ll, h1_end, h1_epi, i, mod1_all_likes,
  N_LEE_SIMS, N_PARAMS, NP_EVAL, NREPS_EVAL, end_params, epi_params,
  pf_params, j, ll_evals, which_params
)
gc()
@

@@

\begin{tabular}{|c|c|c|c|}
\hline
 & \thead{Model~1} & \thead{Model~2} & \thead{Model~3}
\\
\hline
\hline
\multirow{2}{*}{Log-likelihood} &
  $\Sexpr{myround(mod1_ll, 1)}$ &
  $\Sexpr{myround(mod2_ll, 1)}$ &
  $\Sexpr{myround(mod3_spatPomp_res$ll, 1)}$ \\
    & ($\Sexpr{myround(mod1_lee_ll, 1)}$)\footnotemark[1] &
  ($\Sexpr{myround(h2_fit$leeFit_ll, 1)}$) &
  ($\Sexpr{myround(max(lee3_results[["all_ll_evals"]]), 1)}$)\footnotemark[2]
\\
\hline
Number of &
  $\Sexpr{as.character(mod1_n_params)}$ &
  $\Sexpr{as.character(mod2_n_params)}$ &
  $\Sexpr{mod3_spatPomp_n_params}$ \\
 Fit Parameters &
 (\Sexpr{as.character(mod1_lee_n_params)}) &
 ($\Sexpr{as.character(h2_fit$leeFit_n_params)}$) &
 ($\Sexpr{as.character(lee3_results$n_params, 1)}$)
\\
\hline
\multirow{2}{*}{AIC} &
  $\Sexpr{myround(mod1_aic, 1)}$ &
  $\Sexpr{myround(mod2_aic, 1)}$ &
  $\Sexpr{myround(mod3_spatPomp_aic, 1)}$ \\
  & ($\Sexpr{myround(mod1_lee_aic, 1)}$)\footnotemark[1] &
  ($\Sexpr{myround(mod2_lee_aic, 1)}$) &
  ($\Sexpr{myround(mod3_lee_coupled_aic, 1)}$)\footnotemark[2]
\\
\hline
\thead{Log-ARMA(2,1) \\ Log-likelihood} &
  $\Sexpr{myround(ARMA_benchmarks$m1, 1)}$ &
  $\Sexpr{myround(ARMA_benchmarks$m2, 1)}$ &
  $\Sexpr{myround(ARMA_benchmarks$m3, 1)}$
\\
\hline
\end{tabular}
\caption{\label{tab:likes}Log-likelihood values for each models compared to their ARMA benchmarks.
Values in parenthesis are corresponding values using \citet{lee20} parameter estimates.  $\textsuperscript{1}$The reported likelihood is an upper bound of the likelihood of the \citet{lee20} model. $\textsuperscript{2}$\cite{lee20} fit Model~3 to a subset of the data (March 2014 onward, excluding data from Ouest in 2015-2016).
On this subset, their model has a likelihood of $\Sexpr{myround(max(lee3_results$sub_ll_evals), 1)}$.
On this same subset, our model has a likelihood of $\Sexpr{myround(mod3_spatPomp_res$subset_ll, 1)}$.
See Sec.~S5 of the supplement material for more details on estimating the likelihood of the \cite{lee20} models.
}
\end{table}

### 2022-10-17 si/si.Rnw
Updated how Model 1 is replicated.

The model was implemented by a team at Johns Hopkins Bloomberg School of Public Health (hereafter referred to as the Model~1 authors) in the \code{R} programming language using the \code{pomp} package \citep{king16}.
Original source code is publicly available with DOI: 10.5281/zenodo.3360991.
Despite having source code available, no point estimates for model parameters were provided by \cite{lee20}.
According to the supplement materials, this was because model realizations from a single parameter set retained substantial variability, but multiple realizations from a collection of parameter sets resulted in a reasonable visual fit to the data.
We are also inclined to believe that the use of multiple parameter values was in part intended to account for parameter uncertainty (as mentioned in our main text), an effort by the Model~1 authors that we applaud.
We note, however, that simulations from each of the parameter sets were treated with equal importance when being used to diagnose the model fit and make inference on the system.
This practice is problematic given Figures S8 and S9 of the supplement material, which suggest that some parameter sets that were used for inference were several hundred log-likelihood units lower than the best performing sets of parameters.
When accounting for parameter uncertainty, one should instead weigh resulting model projections based on the likelihood of the set of parameters that were used to obtain the model realizations, as was done by \cite{king15}.

Not providing a point estimate for model parameters also has the effect of making model reproduction a more difficult task.
We therefore rely on the range and median values for each of the final model parameters are provided in Tables S10 and S11 of the supplement, and bivariate relationships between some of the fitted parameters (Figures S8 and S9) in order to replicate their results.

We begin by constructing 300 different parameter sets based on the basic summaries and figures provided by the Model~1 authors.
This was primarily done using truncated-bivariate normal distributions, where samples that are drawn outside of the desired range of parameters are re-sampled from the distribution.
Because the likelihood surface is complex and almost certainly includes higher than two-level interactions between parameter values, we then perform iterated filtering for the epidemic phase of the data using the 300 different parameter sets as starting points.

<<Load lee Model1 Params, echo=FALSE>>=
h1_epi <- haiti1()

RL_dir <- paste0("run_level_", run_level, "/")
lee1_epi_mif <- readRDS(file = paste0("../model1/", RL_dir, "lee1_epi_fit.rds"))
@

<<Simulate Lee Model 1, echo=FALSE, cache=TRUE>>=
N_LEE_SIMS <- switch(run_level,  10,  15,   20)

registerDoRNG(18599687)
foreach(
  i = 1:length(lee1_epi_mif),
  .combine = rbind
) %dopar% {
  guess <- lee1_epi_mif$lee1_epi_mif_params[i, ]
  sims <- simulate(h1_epi, nsim = N_LEE_SIMS, format = 'data.frame',
                   params = guess)  # 20 sims for each set of parameters
  sims$param_set <- i
  sims %>%
    dplyr::select(.id, week, cases, param_set)
} -> all_sims

<<PlotEpiDist, echo=FALSE, fig.cap="Compare to Figure S8 of Lee et al. (2020) supplement.", fig.width=7, fig.height=7, fig.align='center'>>=
lee1_epi_evals <- readRDS(file = paste0("../model1/", RL_dir, "lee1_epi_evals.rds"))

epi_results <- as.data.frame(lee1_epi_mif$lee1_epi_mif_params)
epi_results$logLik <- lee1_epi_evals$pfLL

GGally::ggpairs(
  dplyr::filter(epi_results, logLik > -20000),
  columns = c("rho", "tau", "beta1", "nu", "logLik")
)
@

@@

lee1_end_mif <- readRDS(
  file = paste0("../model1/", RL_dir, "lee1_end_fit.rds")
)

@@

Following the model fitting scheme used by the Model~1 authors, we then use these resulting parameters as starting values for the endemic phase, and refit parameters using MIF2.
To do this, we use the same number of particles and MIF iterations that were used by \citet{lee20}.
The resulting bivariate relations between $\rho, \tau, \beta_1$ and $\nu$ are given in Figure~\ref{fig:plotEndParams}.

<<FitEnd, echo=FALSE, message=FALSE, include=FALSE>>=
foreach(
  i = 1:nrow(lee1_end_mif),
  .combine = rbind
) %dopar% {
  r_params <- lee1_end_mif[i, ]

  lee1_epi_mif$start_states %>%
    filter(param_set == i) %>%
    select(-param_set) %>%
    sapply(., median, na.rm = TRUE) -> rinit_parms

  # Set the initial state
  new_rinit <- Csnippet(
    sprintf(
      "
          S = nearbyint(%f);
          E = nearbyint(%f);
          I = nearbyint(%f);
          A = nearbyint(%f);
          R = nearbyint(%f);
          incid = nearbyint(%f);
          foival = %f;
          Str0 = nearbyint(%f);
          Sout = nearbyint(%f);
          Sin = nearbyint(%f);
          ",
      rinit_parms['S'],
      rinit_parms['E'],
      rinit_parms['I'],
      rinit_parms['A'],
      rinit_parms['R'],
      rinit_parms['incid'],
      rinit_parms['foival'],
      rinit_parms['Str0'],
      rinit_parms['Sout'],
      rinit_parms['Sin']
    )
  )

  h1_end <- h1_end %>% pomp(
    rinit = new_rinit,
    statenames = c(
      "S", "E", "I", "A", "R", "incid",
      "foival", "Str0", "Sout", "Sin"
    )
  )

  coef(h1_end) <- r_params

@@

<<plotEndParams, fig.width=7, fig.height=7, fig.align='center', fig.cap="Bivariate relationships between variables after fitting endemic period. Compare to S9 of Lee et al. (2020) supplement.", cache=TRUE, echo=FALSE>>=
lee1_end_evals <- readRDS(file = paste0("../model1/", RL_dir, "lee1_end_evals.rds"))

end_results <- as.data.frame(lee1_end_mif)
end_results$logLik <- lee1_epi_evals$pfLL

GGally::ggpairs(
  filter(end_results, logLik > -20000),
  columns = c("rho", "tau", "beta1", "nu", "logLik")
)
@

Now that we have fit the parameters, we can simulate from the models.
The results of these simulations are shown in Figure~\ref{fig:plotMod1Sims}.
While not a perfect replication of the results given by \citet{lee20sup}, the similarities between the model simulations, and the fact that we based our implementation of Model~1 primarily on the source code that was provided, makes us confident that we have been able to obtain a reliable representation of the model described by \citet{lee20}.

<<plotMod1Sims, echo=FALSE, fig.cap="Simulations from various parameter sets. Simulations from parameter sets with likelihood $<$ -2200 are removed. Compare to Figure S7 of \\cite{lee20sup}.", cache=TRUE, fig.width=6.5, fig.height=3.75>>=

epi_complete <- filter(epi_results, logLik > -20000)
bad_epi_pars <- which(epi_results$logLik < -20000)

quants <- all_sims %>%
  filter(!param_set %in% bad_epi_pars) %>%
  group_by(week) %>%
  summarize(
    q025 = quantile(cases, probs = 0.025, na.rm = TRUE),
    q50  = quantile(cases, probs = 0.500, na.rm = TRUE),
    q975 = quantile(cases, probs = 0.975, na.rm = TRUE)
  ) %>%
  ungroup() %>%
  mutate(date = lubridate::ymd("2010-10-16") + lubridate::weeks(week))

end_complete <- filter(end_results, logLik > -20000)
bad_end_pars <- which(end_results$logLik < -20000)

quants_end <- end_sims %>%
  filter(!param_set %in% bad_end_pars) %>%
  group_by(week) %>%
  summarize(
    q025 = quantile(cases, probs = 0.025, na.rm = TRUE),
    q50  = quantile(cases, probs = 0.500, na.rm = TRUE),
    q975 = quantile(cases, probs = 0.975, na.rm = TRUE)
  ) %>%
  ungroup() %>%
  mutate(date = lubridate::ymd("2010-10-16") + lubridate::weeks(week))

@@

>>>>

The model was implemented by a team at Johns Hopkins Bloomberg School of Public Health (hereafter referred to as the Model~1 authors) in the \code{R} programming language using the \code{pomp} package \citep{king16}.
Original source code is publicly available with DOI: 10.5281/zenodo.3360991.
The final results reported by the Model~1 authors were obtained by using several different parameter sets rather than a single point estimate.
According to the supplement materials, this was because model realizations from a single parameter set retained substantial variability, but multiple realizations from a collection of parameter sets resulted in a reasonable visual fit to the data.
We are also inclined to believe that the use of multiple parameter values was in part intended to account for parameter uncertainty (as mentioned in our main text), an effort by the Model~1 authors that we applaud.
Simulations from each of the parameter sets, however, were treated with equal importance when being used to diagnose the model fit and make inference on the system.
This is problematic given Figures S8 and S9 of the supplement material, which suggest that some parameter sets that were used for inference were several hundred log-likelihood units lower than other parameter sets.
Such a large difference in log-likelihoods is well beyond the threshold of statistical uncertainty determined by Wilks' theorem.

To fully reproduce the results of the Model~1 authors, it is necessary to use the exact same set of model parameters.
Because the parameters used to obtain the result given by \cite{lee20} are not publicly available, we relied on the source code provided by the Model~1 authors to approximately recreate the parameter set.
Due to software updates since the publication of the source code, we were unable to produce the exact same set of parameters.
Still, running the publicly available source code resulted in a set of parameters that are visually similar to those used by the Model~1 authors (See Figures~\ref{fig:PlotEpiDist} and \ref{fig:plotEndParams}).
Furthermore, simulations using the set of parameters produced by the source code appear practically equivalent to those displayed by \cite{lee20} (See Figure~\ref{fig:plotMod1Sims}).

<<Load lee Model1 Params, echo=FALSE>>=
h1_epi <- haiti1()
RL_dir <- paste0("run_level_", run_level, "/")
@

<<Simulate Lee Model 1, echo=FALSE, cache=TRUE>>=
N_LEE_SIMS <- 20

registerDoRNG(18599687)
foreach(
  i = 1:length(h1LeeStartsEpi),
  .combine = rbind
) %dopar% {
  guess <- h1LeeStartsEpi[i, ]
  sims <- simulate(h1_epi, nsim = N_LEE_SIMS, format = 'data.frame',
                   params = guess)  # 20 sims for each set of parameters
  sims$param_set <- i
  sims %>%
    dplyr::select(.id, week, cases, param_set)
} -> all_sims
@

<<PlotEpiDist, echo=FALSE, fig.cap="Compare to Figure S8 of Lee et al. (2020) supplement.", fig.width=7, fig.height=7, fig.align='center'>>=
GGally::ggpairs(
  h1LeeStartsEpi,
  columns = c("rho", "tau", "beta1", "nu", "loglik")
)
@

@@

@@

<<FitEnd, echo=FALSE, message=FALSE, include=FALSE>>=
foreach(
  i = 1:nrow(h1LeeStartsEnd),
  .combine = rbind
) %dopar% {
  r_params <- h1LeeStartsEnd[i, ][names(coef(h1_end))]
  coef(h1_end) <- r_params

@@

<<plotEndParams, fig.width=7, fig.height=7, fig.align='center', fig.cap="Bivariate relationships between variables after fitting endemic period. Compare to S9 of Lee et al. (2020) supplement.", cache=TRUE, echo=FALSE>>=
GGally::ggpairs(
  h1LeeStartsEnd,
  columns = c("rho", "tau", "beta1", "nu", "loglik")
)
@

<<echo=FALSE, include=FALSE, message=FALSE>>=
h1_epi_evals <- readRDS(paste0("../model1/", RL_dir, "lee1_epi_evals.rds"))
h1_end_evals <- readRDS(paste0("../model1/", RL_dir, "lee1_end_evals.rds"))

mod1_all_likes <- dplyr::inner_join(
  x = h1_epi_evals,
  y = h1_end_evals,
  by = "parid"
  ) %>%
  dplyr::mutate(
    joint_ll = pfLL.x + pfLL.y
  )

# Approximate the likelihood of the Lee et al (2020) version of Model 1 by
# using the parameter sets that resulted in the maximum likelihood.
mod1_lee_ll <- max(mod1_all_likes$joint_ll, na.rm = TRUE)
@

<<plotMod1Sims, echo=FALSE, fig.cap=paste0("Simulations using parameter sets that were generated by running source code provided by Lee et al (2020). Compare to Figure S7 of \\cite{lee20sup}. The upper bound for the likelihood of this model is ", round(mod1_lee_ll), "."), cache=TRUE, fig.width=6.5, fig.height=3.75>>=

quants <- all_sims %>%
  filter(!param_set %in% c(1, 2)) %>%  # Remove parameter sets inconsistent with Lee et al (2020).
  group_by(week) %>%
  summarize(
    q025 = quantile(cases, probs = 0.025, na.rm = TRUE),
    q50  = quantile(cases, probs = 0.500, na.rm = TRUE),
    q975 = quantile(cases, probs = 0.975, na.rm = TRUE)
  ) %>%
  ungroup() %>%
  mutate(date = lubridate::ymd("2010-10-16") + lubridate::weeks(week))

quants_end <- end_sims %>%
  group_by(week) %>%
  summarize(
    q025 = quantile(cases, probs = 0.025, na.rm = TRUE),
    q50  = quantile(cases, probs = 0.500, na.rm = TRUE),
    q975 = quantile(cases, probs = 0.975, na.rm = TRUE)
  ) %>%
  ungroup() %>%
  mutate(date = lubridate::ymd("2010-10-16") + lubridate::weeks(week))

@@

To calculate an upper-bound on the likelihood of this model---which uses a large number of parameters that each have a different likelihood value---we estimate the likelihood corresponding to each parameter set using a particle filter, and report the maximum of these likelihoods.
